---
title: "ST558 Homework 2"
author: "Stephanie Shuai, Jiayin Liu"
format: html
editor: visual
---

# Homework 2

## Task 1

1.  Create two vectors

```{r}
pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

2.  Create subject name and assign name to vectors

```{r}
subject_names <- paste("Subject", 1:20, sep = "_")

names(pre) <- subject_names
names(post) <- subject_names

pre
post
```

3.  Calculate change in blood pressure

```{r}
change <- pre - post

change
```

4.  Calculate the average decrease

```{r}
mean(change)
```

5.  Indices of patients with a positive change

```{r}
which(change>0)
```

6.  Subject the vector to keep only positive changes

```{r}
positive_change <- change[which(change>0)]

positive_change
```

7.  Calculate the average decrease for positive changes only

```{r}
mean(positive_change)
```

## Task 2

1.  Create a data frame

```{r}
bp_df <- data.frame(
                 patient = subject_names, 
                 pre_bp = pre,
                 post_bp = post, 
                 diff_bp = pre - post)
```

2.  Return row where diff_bp column is negative

```{r}
bp_df[bp_df$diff_bp<0,]
```

3.  Return row where diff_bp column is negative

```{r}
bp_df$new <- bp_df$post_bp < 120
```

4.  Print the data frame

```{r}
knitr::kable(bp_df)
```

## Task 3

1.  Create a new data frame with this data in it that is similar to the data frame from task 2

```{r}
x<-c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
y<-c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
subject_names2 <- paste("Subject", 1:10, sep = "_")

OthExp_df<-data.frame(pre_bp=x,post_bp=y,patient=subject_names2,diff_bp=x-y)
```

2.  Now create and store a list with two elements

```{r}
Exp_list<-list(treatment=bp_df,placebo=OthExp_df)
```

3.  Access the first list element using three different types of syntax

```{r}
Exp_list[1]
Exp_list[[1]]
Exp_list$treatment
```

4.  In one line, access the placebo data frame, pre_bp column.

```{r}
Exp_list[[2]][[1]]
```

## Task 4

1.  Create a new column in each data frame from above called status

```{r}
Exp_list$treatment$status<-character(20)
Exp_list$placebo$status<-character(10)
```

2.  For the non-placebo data frame (within the list), create a for loop and use if/then/else logic to create the status column’s values.

```{r}
for (i in 1:20) {
  if (Exp_list$treatment$post_bp[i] <= 120) {
    Exp_list$treatment$status[i] <- "optimal"
  } else if (Exp_list$treatment$post_bp[i] > 120 & Exp_list$treatment$post_bp[i] <= 130) {
    Exp_list$treatment$status[i] <- "borderline"
  } else if (Exp_list$treatment$post_bp[i] > 130) {
    Exp_list$treatment$status[i] <- "high"
  } else {
    Exp_list$treatment$status[i] <- 'error'
  }
}
```

3.  Repeat for the placebo data frame (within the list).

```{r}
for (i in 1:10) {
  if (Exp_list$placebo$post_bp[i] <= 120) {
    Exp_list$placebo$status[i] <- "optimal"
  } else if (Exp_list$placebo$post_bp[i] > 120 & Exp_list$placebo$post_bp[i] <= 130) {
    Exp_list$placebo$status[i] <- "borderline"
  } else if (Exp_list$placebo$post_bp[i] > 130) {
    Exp_list$placebo$status[i] <- "high"
  } else {
    Exp_list$placebo$status[i] <- 'error'
  }
}
```

## Task 5

1.  Write a function for data summary

```{r}
summary_bp<-function(data_list,stats='mean'){
  
  my_fun <- get(stats)
  
  treatment_pre  <- my_fun(data_list$treatment$pre_bp)
  treatment_post <- my_fun(data_list$treatment$post_bp)
  treatment_diff <- my_fun(data_list$treatment$diff_bp)
  placebo_pre  <- my_fun(data_list$placebo$pre_bp)
  placebo_post <- my_fun(data_list$placebo$post_bp)
  placebo_diff <- my_fun(data_list$placebo$diff_bp)
  
  out<-c(treatment_pre,treatment_post,treatment_diff,
         placebo_pre,placebo_post,placebo_diff)
  
  names(out)<-c(paste('pre treatment blood pressure',stats),
                paste('post treatment blood pressure' ,stats),
                paste('difference between treatment blood pressure',stats) ,
                paste('pre placebo blood pressure',stats),
                paste('post placebo blood pressure',stats) ,
                paste('difference between placebo blood pressure',stats ))
  return(out)
}


```

2.  Use the function

```{r}
summary_bp(Exp_list)
summary_bp(Exp_list,stats='var')
summary_bp(Exp_list,stats='sd')
summary_bp(Exp_list,stats='min')
summary_bp(Exp_list,stats='max')
```
